\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{amssymb}


\title{Álgebra Lineal}
\author{Héctor Astudillo}
\date{January 2025}

\begin{document}

\maketitle

\section{Espacios Vectoriales}
\subsection*{Vectores}
Sea \( F \) un campo. Un espacio vectorial sobre \( F \) (denotado como \( EV/F \)) es un conjunto \( V \) (cuyos elementos son los vectores que pertenecen al espacio vectorial), equipado con dos operaciones que satisfacen 8 axiomas:

\begin{itemize}
    \item \(EV1\) Commutativity of addition: 
    Let \( u, v \in V \), then \( u + v \in V \).
    \item \(EV2\) Associativity of addition:
    For all \( u, v, w \in V \), we have that \( u + (v + w) = (u + v) + w \).
    \item Commutativity of addition: 
    For all \( u, v \in V \), we have that \( u + v = v + u \).
    \item \(EV3\) Elemento neutro aditivo - Additive identity: 
    Exists \( 0 \in V \) such that \( u + 0 = u \) para todo \( u \in V \).
    \item \(EV4\) Elemento inverso aditivo - Additive inverses:
    For each \( u \in V \), exist a unique element \( -u \in V \) such that \( u + (-u) = 0 \).
    \item \(EV5\) Scalar Product/Multiplication: 
    For all \( c, d \in F \) and \( u \in V \), it satisfies that \( (c \cdot d) \cdot u = c \cdot (d \cdot u) \).
    \item \(EV6\) Neutro/Identidad escalar - Identity element of Scalar Multiplication: 
    For all \( u \in V \), it follows that \( 1 \cdot u = u \), where \( 1 \) is the identity element of the product from the field \( F \).
    \item \(EV7\) Distributivity of scalar over field addition: 
    For all \( c, d \in F \) and \( u \in V \), it holds that \( (c + d) \cdot u = c \cdot u + d \cdot u \).
    \item \(EV8\) Distributivity of scalar over vector addition: 
    Para todo \( c \in F \) y \( u, v \in V \), se cumple \( c \cdot (u + v) = c \cdot u + c \cdot v \).
\end{itemize}

\(F^n\) significa las n-adas de números del campo \(F\)

Por ejemplo, si \(x \in F^n\) entonces hablamos de un vector con \(n\) entradas cuyos coeficientes pertenecen al campo \(F\).
\\
\subsection*{Matrices}
Sean \(m,n \in N\) y \(F\) un campo definimos a \(F^{mxn}\) como el conjunto de todas las matrices de tamaño \(m x n\) con entradas en \(F\)
\\
\\
Remember that:
\\
The sum of matrices is equal entry by entry.
\\
\\
Let us formally define the addition and scalar product on the vectorial space \(F^{mxn}\).
\\

Addition:

Let \(M, N \in F\)...

For every \(i \leq m\) and every \(j \leq m\) we have that \((M + N)_{ij} = M_{ij} + N_{ij}\)
\\

Scalar Product

\(\alpha \in F\), \(m \in F^{mxn}\) such that \(\alpha M \in F^{mxn}\)

Then, \(i \leq m\) and every \(j \leq m\) we have that \((M + N)_{ij} := \alpha \cdot M_{ij}\)
\\
\\
Now we can see that this kind of spaces satisfy the remaining 6 axioms.
\\
\\
Another example... now, we are going to work with functions.

\subsection*{Functions}
Let \(F\) be a field and \(S\) \textbf{a non-empty set}
\\
We define \(F^S\) as \((f | f: S \rightarrow F)\)
\\
In this case, \(S\) is the codomain and \(F\) (the field) is the domain.
\\

Notation:

We can express a function in 2 different ways...
\begin{itemize}
    \item \(f : [a,b] \rightarrow R\)
    \item \(F \in R^{[a,b]}\)
\end{itemize}

Remember the following from Calculus 1:

Let \(f,g \in R^{[a,b]}\) the addition of this functions \(f+g\) maps from \([a,b]\) to \( \mathbb{R}, \forall x \in [a,b]\)

Then, \((f+g)(x) = f(x) + g(x)\)
\\
\\
So it´s the same for our example, now we are going to define the scalar product:

Let \(\alpha \in F\) and \(f \in F^S\)

We want to get \(\alpha \cdot f \in F^S\)

\(\forall t \in S\)

\((\alpha \cdot f)(t) := \alpha \cdot f(t)\)
\\
\\
We already have the addition and scalar product now we can compute the remaining 6 axioms, but there is something else if we want to prove that a functions is equal to another function, we need to prove that the codomain, domain and rule of correspondence it´s the same for both.
\\

Lets shall the \(EV8\)

Let \(\alpha \in F\) and \(f,g \in F^S\)

PD. \(\alpha \cdot (f+g) = \alpha f +\alpha g\)
\\

Domain:

\(dom(\alpha(f+g)) = S\)

\(dom(\alpga f + \alpha g) = S\)
\\

Codomain:

\(cod(\alpha(f+g)) = F\)

\(cod(\alpga f + \alpha g) = F\)
\\

For the rule of correspondence:

Let \(t \in S\)

\((\alpha(f+g))(t) = \alpha \cdot (f+g)(t)\)

\(=\alpha \cdot (f(t) + g(t))\) the field has distributivity

\(= \alpha \cdot f(t) + \alpha \cdot g(t)\)

\(=(\alpha f)(t) + (\alpha g)(t)\)

\(=(\alpha f + \alpha g)(t)\)

\subsection*{Polynomials}

A polynomial with coeficients in \(F\) is:

\[
\sum_{i=0}^{n} a_i x^i = a_0 + a_1 x^1 + a_2x^2 + a_3x^3... +a_nx^n} 
\] 
where \(n \in \mathbb{N} \cup {0}\) and \(\{a_i : 0 \leq i \leq n\} \subsetneq F\)
\\
\\
\(F[x]\) is the collection of all the polynomials with coefficients in \(F\)
\\
\\
Now we need to define the addition and scalar product to shall the remaining 6 axioms. Let´s see how to define formally a polynomial:

Let \(p \in F[x] \rightarrow \exists n \in \mathbf{N}\cup\{0\}\)


\(\qquad \qquad \qquad \quad \exists \{a_i : 0\leq i \leq n\} \subset F\)

Such that 

\[ \(p = \sum_{i=0}^{n} a_ix^i\). \]

The same if we have \(q \in F[x]\)

Let \(q \in F[x] \rightarrow \exists m \in \mathbf{N}\cup\{0\}\)


\(\qquad \qquad \qquad \quad \exists \{b_i : 0\leq i \leq m\} \subset F\)

Such that 

\[ \(q = \sum_{i=0}^{m} b_ix^i\). \]
\\
\textbf{Notice that each polynomial has its own highest degree term and set of coefficients}
\\
The degree term n and m could be any value in Natural field, anyone, so one could be bigger than other, that is the reason why we define \(k\) 

Let \(k := max\{m,n\} \rightarrow \forall n < i \leq k\) \((a_i :=0)\)

\(\qquad \qquad \qquad \qquad \qquad \forall m < i \leq k\) \((b_i :=0)\)

This help us to complete, force the polynomial to have the same degree term (the highest)
\\

Let´s shall and make an abstract definition of addition and scalar product with polynomials and then, prove the remaining 8 axioms.
\\
\\
Addition
\\
We are going to use \(k\) the highest value between m and n that complete the polynomial.

Let \(p, q \in F[X]\)

We define the addition of polynomials as follow:
\[
\(p+ q = \sum_{i=0}^{k} (a_i + b_i) x^i\) \]
\\
\\
Scalar Product

Let \(\alpha \in F\) and \(p \in F[x]\)

Remember how to define formally an polynomial:

\(p \in F(x) \rightarrow \exists n \in \mathbf{N}\cup \{0\}\)

\(\qquad \qquad \quad \exists \{a_i : 0 \leq i \leq m \} \subset F\)
\\
\\
\textbf{Remember this order:}

\textbf{Exist a degree term, the polynomial has a limit, you have to define this.}

\textbf{Then you have to define the coefficient of this polynomial, even define the close interval in the same line, this is the index that can take the coefficients, \(i\)}.
\\
\\
Well I notice that I said "the remaining 6 axioms on each section", which is incorrect, there are 8 axioms cause we define the closure of addition and closure under scalar multiplication. In total are 10 axioms, not 8.
\\
Now let´s shall the \(EV1\) axiom. Commutativity of addition.

\end{document}
